Tree growth, mortality, and climatic stress in west-coast states, USA

### 2025-08-21 Manuscript.Rmd
Fix to mortality calcs, changed bootstrap to sample from the entire pool of plots, not within each stratum.

FIA plots consist of four points where all trees with stem diameters greater than or equal to 12.7 cm (for most species measured at 1.37 m above the ground) were tallied in a 7.32 m radius circular subplot (total area 672.5 m$^2$) with additional trees sampled in an annulus around each subplot (macroplot) around each subplot, with a total radius of 18 m from the center point and a total area of 1012 m$^2$.  We did not include data from FIA microplots.  Bechtold and Scott (2005) provide plot design details and measurement protocols.

Since 2000, FIA plots have been selected via a spatially balanced probability sample (McRoberts 2005, Reams et al., 2005) with a target plot density of one ground plot every 24 km$^2$. Some regions and ownerships intensify the plot density; we excluded these additional plots to simplify stratification estimation, relying on [601] FIA-defined strata and using the density of FIA Phase 1 sampling points within and across strata to determine strata weights.

Plots in the three states are sampled every 10 years with a target of 1/10 of plots sampled in a panel every year (Thompson 2015). Our analysis depends on examining the fate and growth of trees from two sequential visits (referred to as the first and second plot visit), ideally from a complete sample panel remeasure.  We obtained our FIA data from the USDA FIA DataMart web portal (https://research.fs.usda.gov/products/dataandtools/fia-datamart; downloaded May 5, 2025).  They included plot data sampled from 2001 to 2023. Our total sample size was [12500] forested plots (Figure 1).

@@

Our study tracked the fate of `r comma(analysis.stats$tree.num)` trees from `r length(sel.spp)` in `r comma(analysis.stats$plot.num)` plots that were visited twice by FIA crews. The two visits for `r pct.remeas[2]`% of plots were 10 years apart(`r pct.remeas[1]`% were 9 years, `r pct.remeas[3]`% were 11 years). Of the `r length(sel.spp)` species, `r analysis.stats$n.gym.spp` are gymnosperms.  Overall growth rates for species varied between `r min.grow.spp$mean` (*`r min.grow.spp$spp`*) and `r max.grow.spp$mean` (*`r max.grow.spp$spp`*, `r t.ref("tbl_growth")`) cm$^2$/decade.  Decadal mean mortality rates varied between `r min.mort.spp$mean` (*`r min.mort.spp$spp`*) and `r max.mort.spp$mean` (*`r max.mort.spp$spp`*, `r t.ref("tbl_mort")`). Growth and mortality rates differed among states (Tables `r tab_nums("tbl_growth", display = "num")` and `r tab_nums("tbl_mort", display = "num")`).

>>>>

FIA plots consist of four points where all trees with stem diameters greater than or equal to 12.7 cm (for most species measured at 1.37 m above the ground) were tallied in a 7.32 m radius circular subplot (total area 672.5 m$^2$) with additional trees sampled in an anulus around each subplot (macroplot) around each subplot, with a total radius of 18 m from the center point and a total area of 1012 m$^2$.  We did not include data from FIA microplots.  Bechtold and Scott (2005) provide plot design details and measurement protocols.

Since 2000, FIA plots have been selected via a spatially balanced probability sample (McRoberts 2005, Reams et al., 2005) with a target plot density of one ground plot every 24 km$^2$. Some regions and ownerships intensify the plot density; we excluded these additional plots to simplify stratification estimation, relying on FIA-defined strata and using the density of FIA Phase 1 sampling points within and across strata to determine strata weights.

Plots in the three states are sampled every 10 years with a target of 1/10 of plots sampled in a panel every year (Thompson 2015). Our analysis depends on examining the fate and growth of trees from two sequential visits (referred to as the first and second plot visit), ideally from a complete sample panel remeasure.  We obtained our FIA data from the USDA FIA DataMart web portal (https://research.fs.usda.gov/products/dataandtools/fia-datamart; downloaded May 5, 2025).  They included plot data sampled from 2001 to 2023.

@@

Our study tracked the fate of `r comma(analysis.stats$tree.num)` trees from `r length(sel.spp)` species in `r comma(analysis.stats$plot.num)` plots that were visited twice by FIA crews. The two visits for `r pct.remeas[2]`% of plots were 10 years apart (`r pct.remeas[1]`% were 9 years, `r pct.remeas[3]`% were 11 years). Plots were distributed in `r analysis.stats$strat.num` strata.  Of the `r length(sel.spp)` species, `r analysis.stats$n.gym.spp` are gymnosperms.  Mean overall growth rates for species varied between `r min.grow.spp$mean` (*`r min.grow.spp$spp`*) and `r max.grow.spp$mean` (*`r max.grow.spp$spp`*, `r t.ref("tbl_growth")`) cm$^2$/decade.  Decadal mean mortality rates varied between `r min.mort.spp$mean` (*`r min.mort.spp$spp`*) and `r max.mort.spp$mean` (*`r max.mort.spp$spp`*, `r t.ref("tbl_mort")`). Growth and mortality rates differed among states (Tables `r tab_nums("tbl_growth", display = "num")` and `r tab_nums("tbl_mort", display = "num")`).
